### 2018-06-26 latex/supplementaries.tex
Added supplementary legends and methods by Dorothée

%%% TeX-master: "manuscript"

\section{Supplementary Legends}

\subsection{Supplementary figure S1}

\textbf{The structurally unresolved loop structure of NamPRT.} Sequence alignment of NamPRT of different species cropped to the region around the unresolved loop structure. Coloured rectangles indicate the enzymes present in the species besides NamPRT; blue: NNMT; black: NADA and NNMT; yellow: NADA; green: NamPRT only. Major clades are indicated for better orientation. Number of amino acid indicated at the top refer to the human protein.

\subsection{Supplementary figure S2}

\todo{Figure must be made!}\textbf{Distribution of NNMT in birds.} ...

\subsection{Supplementary table S1}

\textbf{Overview of kinetic constants used for the construction of the model.}

\subsection{Supplementary table S2}

\textbf{Query proteins used for Blast searches.}