### 2018-06-26 supplements/KineticsTable/kinetics_table.tex
Fixed kinetics table

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{booktabs,colortbl,tabularx}
\usepackage{longtable}
\usepackage{float}
\usepackage{footmisc}
\usepackage{breqn}
\restylefloat{table}

\usepackage[pdftex,bookmarks=false,
pdftex,
%linkcolor=black,
linkcolor=blue,
citecolor=blue,
colorlinks=true,                %%% hyper-references for pdflatex
urlcolor=blue]{hyperref}

\begin{document}

\section*{Supplementary Table 2}
\addcontentsline{toc}{section}{Supplementary Table 2}
\begin{longtable}{p{3cm}p{2cm}p{3cm}cc}
\multicolumn{5}{l}{\textbf{Overview of kinetic constants used for the
construction of the model.}} \cr\toprule
  \cr\textbf{Enzyme} &
  \textbf{EC\linebreak number}&
  \textbf{Kinetic\linebreak parameter} &
  \textbf{References} &
  \textbf{Rate Law} \\
\cr\hline
\endfirsthead
\multicolumn{5}{c}{Table 1 -- Continued}
\cr\hline
\cr\textbf{Enzyme} &
 \textbf{EC\linebreak number}&
  \textbf{Kinetic\linebreak parameter} &
  \textbf{Reference} &
  \textbf{Rate Law} \\
\cr\hline
\endhead
NADA & \href{http://www.chem.qmul.ac.uk/iubmb/enzyme/EC3/5/1/19.html}{3.5.1.19} &$K_M$:9.6$\mu$M  & \cite{Smith2012} & Product inhibition\\
& & $K_{iP}$:120$\mu$M  & &
    \\
    & & $k_{cat}$:0.65$s^{-1}$  & &
    \\
    \hline
NADS & \href{http://www.chem.qmul.ac.uk/iubmb/enzyme/EC6/3/5/1.html}{6.3.5.1}
&$K_M$:190$\mu$M  & \cite{Yi1972} & HMM\\
    & & $k_{cat}$:21$s^{-1}$  & &
    \\  \hline
NMNAT &
\href{http://www.chem.qmul.ac.uk/iubmb/enzyme/EC2/7/7/1.html}{2.7.7.1}/\href{http://www.chem.qmul.ac.uk/iubmb/enzyme/EC2/7/7/18.html}{2.7.7.18}
&$K_{M_{NaMN}}$:67.7$\mu$M  & \cite{Sorci2007}\footnote{Values for NMNAT1 used}
& Substrate Competition\\
    & & $k_{{cat}_{NaMN}}$:42.9$s^{-1}$  & & \\
    & & $K_{M_{NMN}}$:22.3$\mu$M & & \\
    & & $k_{{cat}_{NMN}}$:53.8$s^{-1}$  & & \\
	& & $K_{M_{NMN}}$:59$\mu$M  & & \\
    & & $k_{{cat}_{NAD}}$:129.1$s^{-1}$  &\cite{Berger2005}\footnote{Keq used
    for calculation of turnover rate of reverse reaction} &
    \\
    & & $K_{M_{NaAD}}$:502$\mu$M  & & \\
    & & $k_{{cat}_{NaAD}}$:103.8$s^{-1}$  &\cite{Berger2005}\footnote{Equilibrium constant used for calculation of turnover rate of reverse reaction} &
    \\ \hline
NMNT & \href{http://www.chem.qmul.ac.uk/iubmb/enzyme/EC2/1/1/1.html}{2.1.1.1}
&$K_M$:400$\mu$M  & \cite{Aksoy1994} & Product inhibition\\
& & $K_{iP}$:60$\mu$M  & &
    \\
    & & $k_{cat}$:8.1$s^{-1}$  & \cite{Alston1988}&
    \\  \hline
NamPRT & \href{http://www.chem.qmul.ac.uk/iubmb/enzyme/EC6/3/5/1.html}{6.3.5.1}
&$K_M$:5nM  & \cite{Burgos2008} & HMM\\
    & & $k_{cat}$:0.0077$s^{-1}$  & &
    \\  \hline
    NAPRT &
       \href{http://www.chem.qmul.ac.uk/iubmb/enzyme/EC2/4/2/11.html}{2.4.2.11}
    &$K_M$:23$\mu$M  & \cite{Hanna1983} & HMM\\
    & & $k_{cat}$:3.3$s^{-1}$  & &
    \\  \hline
    SIRT1 &
    \href{http://www.chem.qmul.ac.uk/iubmb/enzyme/EC3/5/1/index.html}{3.5.1.-}
    &$K_M$:29$\mu$M  & \cite{Borra2004} & Product inhibition\\
    & & $K_{iP}$:60$\mu$M  & &
    \\
    & & $k_{cat}$:0.67$s^{-1}$  & &
    \\  \hline
    NT5 &
    \href{http://www.chem.qmul.ac.uk/iubmb/enzyme/EC3/1/3/5.html}{3.1.3.5}
    &$K_{M_{NaMN}}$:3.5mM  & \cite{PMID26385918} & HMM \\
    & & $k_{{cat}_{NaMN}}$:2.8$s^{-1}$
    & & \\
    & &$K_{M_{NMN}}$: 5mM   &  \\
     & & $k_{{cat}_{NMN}}$:0.5$s^{-1}$
     & & \\  \hline
     PNP &
    \href{http://www.chem.qmul.ac.uk/iubmb/enzyme/EC3/2/4/2.html}{2.4.2.1}
    &$K_M$:1.48mM  & \cite{PMID9030766} & HMM \\
    & & $k_{cat}$:40$s^{-1}$
    & & \\  \hline
    NRK &
    \href{http://www.chem.qmul.ac.uk/iubmb/enzyme/EC2/7/1/173.html}{2.7.1.173}
    &$K_M$:3.4$\mu$M & \cite{PMID19027704} & HMM \\
    & & $k_{cat}$:0.23$s^{-1}$
    & & \\  \hline
  \bottomrule
  \label{tab:kinetic}
\end{longtable}

The total enzyme concentration was set to 10 for all enzymes  except NamPRT and NMNAT, for which the concentration was set to 100. Concentration of potential co-substrate was assumed to be constant
and not-limiting for the reaction. Thus they were  implicitly represented by the
maximal velocities given in the table.

\subsection*{Kinetic Rate Laws}

\subsubsection*{Product Inhibition}
\begin{equation}
v=\cfrac{E_T\cdot k_{cat}\cdot S}{K_M + S + \cfrac{K_M\cdot P}{K_{iP}} }
\end{equation}

\subsubsection*{Henry-Michaelis Menten for irreversible reactions (HMM)}
\begin{equation}
v=\frac{E_T\cdot k_{cat}\cdot S}{K_M + S}
\end{equation}

\subsubsection*{Substrate Competition at NMNAT}
\begin{equation}
v=E_T \cdot \cfrac{\cfrac{k_{{cat}_{A}}\cdot A \cdot
B}{K_{{M}_A}}-\cfrac{k_{{cat}_{P}}\cdot P \cdot
Q}{K_{{M}_P}}}{1+\cfrac{A}{K_{{M}_A}}+\cfrac{B}{K_{{M}_B}}+\cfrac{P}{K_{{M}_P}}+\cfrac{Q}{K_{{M}_Q}}}
\end{equation}

\bibliographystyle{plos2009}
\bibliography{tablelibrary}
\end{document}