### 2018-07-07 latex/acknowledgements.tex
Completely separated content and master file

%%% TeX-master: "manuscript"
% !TeX spellcheck=en_GB

\section*{Acknowledgements}

We thank the Norwegian Research Council for funding (grant no. 250395/F20). We furthermore thank for the computation time provided through UNINETT Sigma2 – the National Infrastructure for High Performance Computing and Data Storage in Norway.