### 2018-07-07 latex/manuscript.tex

% !TeX spellcheck=en_GB

\RequirePackage[l2tabu, orthodox]{nag}
\documentclass[paper=a4, 12pt]{scrartcl}

\usepackage{graphicx}
\usepackage[font=footnotesize]{caption}
\usepackage[utf8]{inputenc}
\usepackage[colorlinks=true,
	linkcolor=black,
	citecolor=black,
	filecolor=black,
	urlcolor=black]{hyperref}
\usepackage{textcomp}
\usepackage[left=1cm,right=5cm,top=2cm,bottom=2cm,nohead,nofoot]{geometry}
\usepackage[textsize=scriptsize]{todonotes}

\setlength{\marginparwidth}{4cm}

\frenchspacing
\graphicspath{ {../} }

\begin{document}

\input{header}

\input{introduction}

\input{results}

\input{discussion}

\input{methods}

\bibliographystyle{plain}
\bibliography{library}

\input{acknowledgements}

\input{figurelegends}

\input{supplementaries}

\end{document}