### 2018-07-07 latex/figurelegends.tex
Added Fig S2

%%% TeX-master: "manuscript"
% !TeX spellcheck=en_GB

\newpage

\section*{Figure Legends}

\subsection*{Figure 1}

\begin{figure}[ht]
  \centering
  \includegraphics{fig_1_pathway_overview}
  \caption{\textbf{Schematic overview of NAD biosynthesis pathways} NAD can be synthesised from tryptophan (Trp), nicotinamide (Nam), nicotinic acid (NA), and, to a lesser extend, nicotinamide ribose (NR). Nam is the main precursor in human and also the product of NAD-consuming signalling reactions by enzymes such as sirtuins (NAD-dependent histone deacetylases) or PARPs (poly-ADP-ribosylases). For the recycling of Nam, two different pathways exist. The pathway found in yeast and many bacteria starts with the deamination of Nam by Nam deamidase (NADA). The other three enzymes comprise the Preiss-Handler pathway that also exists in vertebrates. The pathway found in vertebrates directly converts Nam into the corresponding mononucleotide (NMN) by the Nam phosphoribosyltransferase (NamPRT). The Nam N-methyltransferase (NNMT) degrades Nam to methyl-Nam (MNam), which is in mammals excreted with the urine.}
  \label{fig:pathway_overview}
\end{figure}

\newpage

\subsection*{Figure 2}

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.75\textwidth]{fig_2_phylo_distribution}
  \caption{\textbf{Phylogenetic distribution of NADA, NNMT, and NamPRT and their relation to the number of NAD consumers.} A) Distribution of NADA, NNMT, and NamPRT in selected major taxa. NADA is dominant in Bacteria, Fungi, and Plants (Viridiplantae), whereas NamPRT together with NNMT is dominant in Metazoa. Numbers at the pie charts show, how many species of the taxon possess the respective enzyme combination indicated by the colour, which is explained in the lower right of the figure. Below the taxon name, the number of species in that taxon is given. B) Common tree of selected taxa within the Metazoa, including 334 species. The pie charts indicate the distribution of species within the respective taxon that encode the enzyme combination indicated by the different colours. The size of the pie charts is proportional to the logarithm of the number of species analysed in the particular taxon. The numbers below the taxon names indicate the average number of NAD-consuming enzyme families found in all sub-taxa. The branch length is arbitrary.}
  \label{fig:phylo_distribution}
\end{figure}

\newpage

\subsection*{Figure 3}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig_3_NNMT_NAD_flux}
  \caption{\textbf{NNMT enables high NAD consumption flux.} We used a dynamic model of NAD biosynthesis and consumption (for details, see experimental procedures) to simulate NAD consumption flux. The amount of NMNAT and NamPRT used in the simulations where adjusted such that the free NAD concentrations were in the range reported in the literature. All other enzyme concentrations were set equal. Details are given in supplementary table~S2. In the absence of NNMT (blue lines), steady state NAD consumption rates (A) are higher despite reduced NAD concentrations (B). Increasing the amount of NamPRT in the simulation fourfold (blue dotted lines) partially compensates for the decreased NAD steady state concentration caused by Nam degradation through NNMT.}
  \label{fig:NNMT_NAD_flux}
\end{figure}

\newpage

\subsection*{Figure 4}

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\textwidth]{fig_4_NamPRT_affinity_Nam}
  \caption{\textbf{Role of NamPRT affinity for Nam.} Using the dynamic model of NAD biosynthesis and consumption we simulated the affect of different Michaelis-Menten constants ($K_{M}$) of NamPRT for Nam on steady state NAD consumption flux and NAD concentration at different cell division rates. All the parameters were equal to those used for the simulations in figure~\ref{fig:NNMT_NAD_flux}. In the absence of NNMT (A and B), the $K_{M}$ of NamPRT has little influence on NAD consumption and concentration, but both are changing with cell division rates. In the presence of NNMT (C and D), decreasing $K_{M}$ of NamPRT enables increasing NAD consumption flux and NAD concentration. NNMT furthermore makes both, NAD consumption flux and concentration, almost independent of cell division rates. Comparing the situation with and without NNMT (E and F) at different NamPRT $K_{M}$ reveals that at low $K_{M}$ (dashed lines) and high cell division rates NNMT no longer enables higher NAD consumption rates compared to NamPRT alone (green line and dashed black line).}
  \label{fig:NamPRT_affinity_Nam}
\end{figure}

\newpage

\subsection*{Figure 5}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig_5_optimal_substr_affinities}
  \caption{\textbf{The substrate affinities of human NNMT and NamPRT are optimal.} Using the same model as in figures~\ref{fig:NNMT_NAD_flux} and~\ref{fig:NamPRT_affinity_Nam}, we simulated the impact of changes in the $K_{M}$ for both NamPRT and NNMT on NAD consumption rates (A) and NAD concentration (B). Both are increasing with decreasing $K_{M}$ of NamPRT but increasing $K_{M}$ of NNMT. The affinities reported for human enzymes (indicated by a black asterisk) appear to be close to optimal, as further improvements would have little or no effect on NAD consumption or concentration.}
  \label{fig:optimal_substr_affinities}
\end{figure}

\newpage

\subsection*{Figure 6}

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.75\textwidth]{fig_6_unresolved_loop}
  \caption{\textbf{The function of the structurally unresolved loop structure of NamPRT.} Most Deuterostomia that encode NNMT show a sequence insertion in the N-terminal region of NamPRT that has been revealed by multiple sequence alignment of NamPRT from different species (A). Coloured circles indicate the enzymes present in the respective species blue: NamPRT and NNMT; black: NamPRT, NADA and NNMT; yellow: NamPRT and NADA. For a more comprehensive alignment, please see supplementary figure~S1. The visualisation of human NamPRT (B) is based on a structure prediction of SWISS-MODEL \cite{Arnold2006,Biasini2014} using the model 2H3D of the human NamPRT as template \cite{Wang2006}. The inserted region is not resolved in any of currently available crystal structures of NamPRT and thus appears to be a flexible loop structure at the surface of the NamPRT dimer, coloured in red. Immunofluorescence images show that the localisation of the FLAG-tagged mutant protein lacking the unresolved loop (C) is not changed compared to FLAG-tagged human wildtype NamPRT. Both show a heterogeneous nuclear cytosolic localisation in HeLa S3 cells. \textit{In vitro} measurements using recombinant protein show that the mutant NamPRT has a lower activity than the wildtype enzyme (D) and is not activated by ATP (E). Bars in D) and E) with different letters indicate significant difference of measured values as estimated using a T test assuming independent samples and significance at $p < 0.05$.}
  \label{fig:unresolved_loop}
\end{figure}

\newpage

\subsection*{Figure 7}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig_7_NNMT_comp_advantage}
  \caption{\textbf{NNMT provides a competitive advantage and makes NADA obsolete.} To simulate competition for common resources, we created a two compartment model where one compartment contained NADA but no NamPRT and the other compartment contained NamPRT either with or without NNMT but no NADA. NADA and NamPRT were simulated to be present at equal amounts. In the absence of NNMT (A and B) the compartment containing NADA has slightly lower NAD consumption rates (A) but much higher steady state NAD concentrations (B). In the presence of NNMT, however, both NAD consumption (C) and NAD concentration (D) are lower in the NADA compartment. This effect is dependent on a low NamPRT $K_{M}$.}
  \label{fig:NNMT_comp_advantage}
\end{figure}

\newpage

\subsection*{Figure 8}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig_8_evo_events}
  \caption{\textbf{Schematic representation of evolutionary events in the NAD pathway} Based on the phylogenetic analysis presented here and earlier work \cite{Lau2010} we summarized and indicated important events in the evolution of NAD metabolism in Metazoa. Events in italics are from \cite{Lau2010}. All other events were found in this work.}
  \label{fig:evo_events}
\end{figure}

\todo{What about the positive selection in NNMT?}

\newpage