### 2018-11-15 latex/acknowledgements.tex
added additional authors and affiliations, revised introduction

%%% TeX-master: "manuscript"
% !TeX spellcheck=en_GB

\section*{Acknowledgements}

We thank the Norwegian Research Council for funding (grant no. 250395/F20 and grant no. 226244/F50). We furthermore thank for the computation time provided through UNINETT Sigma2 – the National Infrastructure for High Performance Computing and Data Storage in Norway.

\section*{Author contribution}
MB and TG performed the phylogenetic analysis, IH performed the mathematical modelling, DH and MN performed the experiments, IR performed the MD analyses guided by AS, MZ and IH were supervising and guiding the investigations.
All authors contributed to the manuscript preparation..